### 2019-12-21 preamble/font/fontspec/cm-unicode-concrete.tex
Поддержка руских метаданных

%% Computer Modern Unicode fonts
%% url: http://canopus.iacp.dvo.ru/~panov/cm-unicode/
%% gentoo: emerge media-fonts/cm-unicode-otf

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{CMU Concrete}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{CMU Sans Serif}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{CMU Typewriter Text}

\setmainfont[Ligatures=TeX]{CMU Concrete}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{CMU Sans Serif}
\setmonofont[Scale=MatchLowercase]{CMU Typewriter Text}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: