### 2019-12-21 preamble/font/fontspec/liberation.tex

%% A GPL-2 Helvetica/Times/Courier replacement TrueType font set, courtesy of Red Hat
%% url: https://fedorahosted.org/liberation-fonts
%% gentoo: emerge media-fonts/liberation-fonts

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{Liberation Serif}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{Liberation Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{Liberation Mono}

\setromanfont[Ligatures=TeX]{Liberation Serif}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{Liberation Sans}
\setmonofont[Scale=MatchLowercase]{Liberation Mono}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: