### 2019-12-21 preamble/font/fontspec/oldfonts-academy.tex

%% A Academy Old (Академическая) font from OldFonts bundle
%% url: http://lizard.phys.msu.su/home/compu_sci/
%% gentoo: emerge media-fonts/oldfonts-ttf

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{Academy Old}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{Liberation Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{Liberation Mono}

\setromanfont[Ligatures=TeX]{Academy Old}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{Liberation Sans}
\setmonofont[Scale=MatchLowercase]{Liberation Mono}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: