### 2020-09-24 template/preamble/font/fontspec/dejavu.tex
feat: move all in template directory

%% DejaVu fonts, bitstream vera with ISO-8859-2 characters
%% url: http://dejavu.sourceforge.net/
%% gentoo: emerge media-fonts/dejavu

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{DejaVu Serif}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchUppercase]{DejaVu Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchUppercase]{DejaVu Sans Mono}

\setmainfont[Ligatures=TeX]{DejaVu Serif}
\setsansfont[Ligatures=TeX,Scale=MatchUppercase]{DejaVu Sans}
\setmonofont[Scale=MatchLowercase,Scale=MatchUppercase]{DejaVu Sans Mono}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: