### 2020-09-24 template/preamble/font/fontspec/libertine.tex

%% OpenType fonts from the Linux Libertine Open Fonts Project
%% url: http://linuxlibertine.sourceforge.net/
%% gentoo: emerge media-fonts/libertine

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{Linux Libertine O}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{Linux Biolinum O}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{Linux Libertine Mono O}

\setromanfont[Ligatures=TeX]{Linux Libertine O}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{Linux Biolinum O}
\setmonofont[Scale=MatchLowercase]{Linux Libertine Mono O}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: