### 2020-09-24 template/preamble/font/fontspec/paratype-pt.tex

%% ParaType font collection for languages of Russia
%% url: http://www.paratype.com/public
%% gentoo: emerge media-fonts/paratype

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{PT Serif}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{PT Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{PT Mono}

\setromanfont[Ligatures=TeX]{PT Serif}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{PT Sans}
\setmonofont[Scale=MatchLowercase]{PT Mono}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: