### 2024-07-21 template/preamble/font/fontspec/paratype-pt+noto.tex
feat(class): switch to new tex-class

%% ParaType PT font collection for languages of Russia
%% url: http://www.paratype.com/public
%% url: https://company.paratype.com/pt-sans-pt-serif
%% url: https://en.wikipedia.org/wiki/PT_Fonts
%% gentoo: emerge media-fonts/paratype

\setromanfont[Ligatures=TeX]{PT Serif}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{PT Sans}
\setmonofont[Scale=MatchLowercase,Scale=0.8]{Noto Sans Mono}

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{PT Serif}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{PT Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase,Scale=0.8]{Noto Sans Mono}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: