### 2024-07-21 template/preamble/font/fontspec/paratype-pt-astra.tex

%% ParaType PT Astra font collection for languages of Russia
%% url: https://www.paratype.ru/cinfo/news.asp?NewsId=3469
%% url: https://en.wikipedia.org/wiki/PT_Fonts
%% gentoo: emerge media-fonts/paratype-astra

\setromanfont[Ligatures=TeX]{PT Astra Serif}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{PT Astra Sans}
\setmonofont[Scale=MatchLowercase]{PT Mono}

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{PT Astra Serif}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{PT Astra Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{PT Mono}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: