### 2024-07-21 template/preamble/font/fontspec/plex.tex

%% IBM Plex fonts
%% url: https://github.com/IBM/plex
%% gentoo:
%% emerge media-fonts/ibm-plex
%% emerge media-fonts/stix-fonts

\setmainfont[Ligatures={Common,TeX},Scale=0.94]{IBM Plex Serif}
\setromanfont[Ligatures={Common,TeX},Scale=0.94]{IBM Plex Serif}
\setsansfont[Ligatures={Common,TeX},Scale=MatchLowercase,Scale=0.94]{IBM Plex Sans}
\setmonofont[Scale=MatchLowercase,Scale=0.94,FakeStretch=0.9]{IBM Plex Mono}

\addfontfeatures{Fractions=On}

\makeatletter
\AtBeginDocument{%
  \@ifpackageloaded{unicode-math}{%
    \setmathfont{STIX Two Math}
  }{%
  }
}
\makeatother

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: