### 2024-07-21 template/preamble/font/fontspec/stix2.tex

%% Scientific and Technical Information Exchange (STIX) fonts
%% url: https://www.stixfonts.org/
%% url: https://github.com/stipub/stixfonts
%% gentoo:
%% emerge media-fonts/stix-fonts
%% emerge media-fonts/source-code-pro
%% emerge media-fonts/source-sans

\setmainfont[Ligatures={Common,TeX}]{STIX Two Text}
\setromanfont[Ligatures={Common,TeX}]{STIX Two Text}
\setsansfont[Ligatures={Common,TeX},Scale=MatchLowercase]{Source Sans 3}
\setmonofont[Scale=MatchLowercase,Scale=0.92]{Source Code Pro}

\makeatletter
\AtBeginDocument{%
  \@ifpackageloaded{unicode-math}{%
    \setmathfont{STIX Two Math}
  }{%
  }
}
\makeatother

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: