### 2024-07-21 template/preamble/font/fontspec/cm-unicode.tex

%% Computer Modern Unicode fonts
%% url: http://canopus.iacp.dvo.ru/~panov/cm-unicode/
%% gentoo: emerge media-fonts/cm-unicode-otf

\setmainfont[Ligatures=TeX,Scale=MatchLowercase]{CMU Serif}
\setsansfont[Ligatures=TeX,Scale=MatchLowercase]{CMU Sans Serif}
\setmonofont[Scale=MatchLowercase]{CMU Typewriter Text}

% CMU Serif
\newfontfamily\cyrillicfont%
[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]%
{CMU Serif}

% CMU Sans Serif
\newfontfamily\cyrillicfontsf%
[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]%
{CMU Sans Serif}

% CMU Typewriter Text
\newfontfamily\cyrillicfonttt%
[Script=Cyrillic,Scale=MatchLowercase]%
{CMU Typewriter Text}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: