### 2024-07-21 template/preamble/font/fontspec/libertinus.tex

%% The font family Libertinus is derived from the Linux Libertine and enhanced with a math font by Khaled Hosny.
%% url: ctan.org/fonts/libertinus

\newfontfamily\cyrillicfont[Script=Cyrillic,Ligatures=TeX]{Libertinus Serif}
\newfontfamily\cyrillicfontsf[Script=Cyrillic,Ligatures=TeX,Scale=MatchLowercase]{Libertinus Sans}
\newfontfamily\cyrillicfonttt[Script=Cyrillic,Scale=MatchLowercase]{Libertinus Mono}

\usepackage{libertinus-otf}

%%% Local Variables:
%%% mode: latex
%%% coding: utf-8-unix
%%% End: